Figures for Paper

### 2020-11-25 paper_figures.Rmd
Continuing to fill out the Rmd

ipsc_splicing %>% 
  mutate(junction_name = glue::glue("{gene_name} - {paste_into_igv_junction}")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = ifelse(log10_test_stat == 16 & deltaPSI > 0 , junction_name,"")) %>%
  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
                  min.segment.length = unit(0, 'lines')) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  guides(alpha = FALSE, size = FALSE) + 
  scale_fill_manual(values = junc_colors) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab("-Log 10 Test Statistic") + 
  theme(text = element_text(size = 18))
```

@@

```{r}
rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

@@

clean_data_table %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1)) %>% 
    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn_2_cryptic_psi_leaf_plusone",
                      y = "log10_unc13a_cryptic_leaf_psi_plusone",
                      color = 'call',fill = 'call',                      
                      add = "reg.line") +
    ylab("Log10 UNC13A Cryptic PSI + 1") + 
    xlab("Log10 STMN2 Cryptic PSI + 1 ") +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) + 
    scale_fill_manual(values = c("#88CCEE","#44AA99","#105e2a"))

@@

print(pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI"))
```

@@

```{r}
overall_fisher = clean_data_table %>%

  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  dplyr::select(participant_id,unc13a_detected,call) %>% 
  unique() %>% 
  group_by(call) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
  dplyr::select(call,n_sample,n_sample_detected) %>% 
  unique()

@@

overall_fisher %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))

@@

```{r}
detection_table = clean_data_table %>% 
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
  add_count(name = "genotype_detected") %>% 
  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
  unique() %>% 
  pivot_wider(names_from = "unc13a_detected",
              values_from = "genotype_detected") %>% 
  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
  ungroup() %>% 
  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))

count <- 0
breaks_fun <- function(x) {
  count <<- count + 1L
  switch(
    count,
    detection_table$detection_name[1:3],
    detection_table$detection_name[4:6],
    detection_table$detection_name[7:9],
  )
}

detection_table %>% 
  ggplot(aes(x = 1, y = detection_rate,fill = call)) + 
  geom_col(show.legend = F,position = 'dodge2') + 
  facet_wrap(~stmn2_psi_groups) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) + 
  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
  ylab("Percent of Samples \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18))  
  # xlab("N Tissues") +
  # scale_x_continuous(labels = breaks_fun, breaks = c(0.75, 1, 1.25), limits = c(0, NA)) 
```

>>>>

ipsc_splicing %>% 
  mutate(junction_name = glue::glue("{gene_name} - {paste_into_igv_junction}")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ "")) %>% 
  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
                  min.segment.length = unit(0, 'lines')) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  guides(alpha = FALSE, size = FALSE) + 
  scale_fill_manual(values = junc_colors) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab("-Log 10 Test Statistic") + 
  theme(text = element_text(size = 18))
```

@@

We have 2 PSI's for the UNC13A cryptic, one which includes both the short and
long form of the cryptic, and one which does not. While we're not mentioning in 
figure 1, given that the longer form of the cryptic appears in control cerebellum, 
one could argue that that junction should not be included in the PSI calculation
here for UNC13A cryptic per se. In practice, this makes very little difference in the
conclusions made on the cell lines, so I've included both for completeness.

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The NB cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression. 
```{r}
rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

 rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE) 
```

@@

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1)) %>% 
    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
    group_by(call) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn_2_cryptic_psi_leaf_plusone",
                      y = "log10_unc13a_cryptic_leaf_psi_plusone",
                      color = 'call',fill = 'call',                      
                      add = "reg.line") +
    ylab("Log10 UNC13A Cryptic PSI + 1") + 
    xlab("Log10 STMN2 Cryptic PSI + 1 ") +
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) + 
    scale_fill_manual(values = c("#88CCEE","#44AA99","#105e2a"))

@@

the_pretty_plot = pretty_effect_plot(another_model2,p_value_cutoff = 1,"UNC13A Cryptic PSI") + ggpubr::theme_pubr()
print(the_pretty_plot)
```

@@

```{r}
overall_fisher = clean_data_table %>%
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  dplyr::select(participant_id,unc13a_detected,call) %>% 
  unique() %>% 
  group_by(call) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
  dplyr::select(call,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(n_non_detected = n_sample - n_sample_detected) %>% 
  dplyr::select(-n_sample)

@@

overall_fisher %>% 
  mutate(n_sample = n_non_detected + n_sample_detected) %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(labels = scales::percent) + 
  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))

@@

Only looking at disease relevant tissue and in patients
```{r}
detection_table = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                        stmn2_psi_groups == 2 ~ "High STMN2",
                                        TRUE ~ stmn2_psi_groups)) %>%
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
  add_count(name = "genotype_detected") %>% 
  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
  unique() %>% 
  pivot_wider(names_from = "unc13a_detected",
              values_from = "genotype_detected") %>% 
  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
  ungroup() %>% 
  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))

detection_table %>% 
  ggplot(aes(x = call, y = detection_rate,fill = call)) + 
  geom_col(show.legend = F,position = 'dodge2') + 
  facet_wrap(~stmn2_psi_groups) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) + 
  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
  ylab("Percent of Samples \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18))  +
  ggpubr::theme_pubr() + 
  geom_text(aes(label = total_tissue,y = 0),vjust = 1,size = 6) + 
  theme(text = element_text(size = 18)) +
  ggpubr::theme_pubr() +
  xlab("N Tissues") 
```

UNC13A cryptic PSI

```{r}
clean_data_table %>%   
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(stmn2_annotated > 10) %>% 
    filter(cryptic_psi > 0) %>% 
    ggbarplot( x = "call", 
               y = "cryptic_psi", 
               color = 'call',
              position = position_dodge(0.8),
              add = c("mean_se","jitter")) + 
    ylab("UNC13A Cryptic PSI") + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
    theme(text = element_text(size = 20)) + 
    stat_compare_means(aes(group = call), label = "p.format") + 
    stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif")
```

```{r}
clean_data_table %>%   
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                        stmn2_psi_groups == 2 ~ "High STMN2",
                                        TRUE ~ stmn2_psi_groups)) %>%
    filter(cryptic_psi > 0) %>%
    ggbarplot( x = "call", 
               y = "cryptic_psi", 
               color = 'call',
               facet.by = 'stmn2_psi_groups',
              position = position_dodge(0.8),
              add = c("mean_se","jitter")) + 
    ylab("UNC13A Cryptic PSI") + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#105e2a")) +
    theme(text = element_text(size = 20)) + 
    stat_compare_means(aes(group = call), label = "p.format") + 
    stat_compare_means(aes(group = call),comparisons = genotype_comparisons,label = "p.signif")  + 
  xlab("")
```
